Draft-assinment1

### 2017-02-12 Draft-xuezhou.Rmd
xiang

library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)

@@

variable = select(data, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
                  NPT4_PUB,
                  NPT4_PRIV,
                  UG,
                  UG_NRA,
                  UG_UNKN,
                  UG_WHITENH,
                  UG_BLACKNH,
                  UG_API,
                  UG_AIANOLD,
                  UG_HISPOLD,
                  DEP_STAT_PCT_IND,
                  PAR_ED_PCT_1STGEN,
                  MD_FAMINC,
                  MEDIAN_HH_INC,
                  POVERTY_RATE,
                  PCTPELL,
                  PCTFLOAN,
                  GRAD_DEBT_MDN,
                  WDRAW_DEBT_MDN,
                  C150_4,
                  MD_EARN_WNE_P6,
                  RPY_3YR_RT_SUPP,
                  CDR3)
```

@@

#combine
data_new <- cbind(variable_num,variable$preF, variable$ctrF,variable$STABBR)
```

@@

| --------------  | ----------------- | --------------------------------- | -------- |
| NPT4_COMBINE    | Numeric           |  Continuous                       |  Cost    |
| CDR3             | Numeric           |  Continuous                       | n        |
| variable$qreF    | Factor            |  Categorical                      |  n       |
| variable$ctrF    | Factor            |  Categorical                      |   n      |
| variable$STABBR  | Factor            |  Categorical                      |   n      |

>>>>

library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(plyr)

@@

variable = select(data, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
                  NPT4_PUB,
                  NPT4_PRIV,
                  UG,
                  UG_NRA,
                  UG_UNKN,
                  UG_WHITENH,
                  UG_BLACKNH,
                  UG_API,
                  UG_AIANOLD,
                  UG_HISPOLD,
                  DEP_STAT_PCT_IND,
                  PAR_ED_PCT_1STGEN,
                  MD_FAMINC,
                  MEDIAN_HH_INC,
                  POVERTY_RATE,
                  PCTPELL,
                  PCTFLOAN,
                  GRAD_DEBT_MDN,
                  WDRAW_DEBT_MDN,
                  APPL_SCH_PCT_GE2,
                  C150_4,
                  MD_EARN_WNE_P6,
                  MD_EARN_WNE_P10,
                  RPY_3YR_RT_SUPP,
                  CDR3)

@@

#combine
data_new <- cbind(variable_num,variable$preF, variable$ctrF,variable$STABBR)

@@

| Variable Name    | Data Type         | Categorical or Continuous         | Group       |
| --------------  | ----------------- | --------------------------------- | -------- |
| NPT4_COMBINE    | Numeric           |  Continuous                       |  Cost    |
| CDR3             | Numeric           |  Continuous                       | n        |
| variable$qreF    | Factor            |  Categorical                      |  n       |
| variable$ctrF    | Factor            |  Categorical                      |   n      |
| variable$STABBR  | Factor            |  Categorical                      |   n      |
| DEP_STAT_PCT_IND | Numeric           |  Continuous                       |     Student    |
| PAR_ED_PCT_1STGEN| Numeric           |  Continuous                       |      Student   |
| MD_FAMINC        | Numeric           |  Continuous                       |       Student  |
| PCTPELL          | Numeric           |  Continuous                       |Aid         |
| PCTFLOAN         | Numeric           |  Continuous                       | Aid        |
| GRAD_DEBT_MDN    | Numeric           |  Continuous                       |  Aid      |
| APPL_SCH_PCT_GE2 | Numeric           |  Continuous                       | Student        |
| WDRAW_DEBT_MDN   | Numeric           |  Continuous                       |    Aid     |
| MD_EARN_WNE_P6   | Numeric           |  Continuous                       |     Earning    |
| MD_EARN_WNE_P10  | Numeric           |  Continuous                       |       Earning  |
| RPY_3YR_RT_SUPP  | Numeric           |  Continuous                       |      Repayment   |
| CDR3             | Numeric           |  Continuous                       | Repayment        |
| variable$qreF    | Factor            |  Categorical                      |  School       |
| variable$ctrF    | Factor            |  Categorical                      |   School      |
| variable$STABBR  | Factor            |  Categorical                      |   School      |

##School

###1. State
```{r, message=FALSE}
state_freq = count(data_clean,"variable$STABBR")[order(-count(data_clean,'variable$STABBR')$freq),]
state_freq[1:15,]
data_clean$STABBR1 <- with(data_clean, reorder(variable$STABBR, variable$STABBR, function(x) -length(x)))
data_clean %>% ggplot(aes(x= STABBR1)) +
  geom_histogram(stat = "count") 
```

###2. Predominant Degree
```{r, message=FALSE}
prop.table(table(data_clean$`variable$preF`)) * 100
data_clean %>% ggplot(aes(x=variable$`preF`, fill = 'blue')) + 
  geom_histogram(stat = "count") +
  xlab("Degree") 
```

###3. Control of Institution

```{r, message=FALSE}
prop.table(table(data_clean$`variable$ctrF`)) * 100
data_clean %>% ggplot(aes(x=variable$`ctrF`, fill = 'blue')) + 
  geom_histogram(stat = "count") +
  xlab("Degree")
```

##Student
###3. How many application do students submit?
```{r, message=FALSE}
college_app_2_clean = data_clean$APPL_SCH_PCT_GE2[!is.na(data_clean$APPL_SCH_PCT_GE2)]*100
summary(college_app_2_clean)

college_app_2_clean %>% as.data.frame() %>%
    mutate(college_app_1 = 100-college_app_2_clean)%>%
    {.}  -> college_app
colnames(college_app)[1] <- "two_or_greater"

college_app %>% gather(key = num_app, value = perc_student) %>%
{.} -> college_app

college_app %>%
    ggplot(aes(factor(num_app), perc_student)) +
    geom_boxplot() 
```

##Cost
###1.Net Price

##Aid

###1. The median debt for students who have completed

###2. The median debt for students who have not completed
```{r, message=FALSE}
median_debt_non_null = data_clean$WDRAW_DEBT_MDN[!is.na(data_clean$WDRAW_DEBT_MDN)]
summary(median_debt_non_null)
sd(median_debt_non_null)

#plot a ggplot to better illustrate
data_clean %>% ggplot(aes(x = WDRAW_DEBT_MDN)) +
                geom_histogram(binwidth = 450) #warning?

ggplot(data=data_clean,
       aes(x=WDRAW_DEBT_MDN,
           fill = I('red'))) +
    geom_density(alpha=0.5) +
    scale_x_continuous("Median Debt for Students Who Have Completed")
```

##Earning

###. Earning after graduation
```{r,message=FALSE}
earning = select(data_clean, MD_EARN_WNE_P6, MD_EARN_WNE_P10 )
earning_clean = na.omit(earning)
earning_clean %>%
    mutate(salary_increase = MD_EARN_WNE_P10-MD_EARN_WNE_P6) %>%
    {.} -> earning_clean
summary(earning_clean)
apply(earning_clean, 2, sd) #dispersion increases

earning_clean %>% gather(key = years_from_entry, value = salary) %>%
    {.} -> earning_clean1

earning_clean1[earning_clean1$years_from_entry != 'salary_increase',] %>% 
    ggplot(aes(x = salary, fill =years_from_entry )) +
    geom_density(alpha = 0.5) + 
    xlim(0, 150000)
```

##Repayment

###1. 3-year Repayment rate
```{r,message=FALSE}
repayment_clean = na.omit(data_clean$RPY_3YR_RT_SUPP)
summary(repayment_clean)
qplot(repayment_clean, geom = "histogram")
```

###2. 3-year Default rate
```{r,message=FALSE}
default_clean = na.omit(data_clean$CDR3)
qplot(default_clean, geom = "histogram")

summary(data_clean$RPY_3YR_RT_SUPP + data_clean$CDR3)
```
